Digital economy in the UK: a multi-scalar story of the diffusion of web technologies

### 2023-11-24 paper/submission1/tranos2023.qmd
CAT OA regions

<!-- geographers lack of interest + motivation -->
Geographers were always interested in how new technologies and
innovations diffuse across space and, importantly, how such
spatio-temporal processes can be modelled. After all, diffusion together 
with invention and innovation are considered the pillars of 
technological change [@das2022diffusion]. The seminal contribution of
@hagerstrand1968innovation is illustrative of this early interest.
However, the torch of exploring and modelling such processes had been
passed to other disciplines such as economics, business studies and
sociology well before the 'cultural turn' of economic geography
[@perkins2005international]. A potential explanation of the lack of 
geographical studies exploring the diffusion of new and, more 
specifically, digital technologies across *both* space and time can be
attributed to the scarcity of relevant and granular enough data. As 
@zook2022mapping highlight, digital activities are hardly ever captured 
in official data.

<!-- contribution -->
This paper offers such a contribution: a geographical study illustrating 
how a new technology that is the web diffused over space and time 
in the UK at a high level of spatial granularity during the $1996-2012$ 
period. It does so by employing a novel source of big data which 
captures the active engagement with web technologies during that period.
By addressing this empirical question this paper exemplifies how the 
combination of data sources which escape the traditional social science
domain and adequate research methods can offer new contributions 
regarding our understanding of how new technologies diffuse.

<!-- motivation -->
The motivation for this paper lies in the fact that there are various
stakeholders who are interested in knowing how new digital technologies
diffused over space and time and use this knowledge to make predictions
regarding the diffusion of related *future* technologies.
As per @leibowicz2016representing, historical studies agree on the fact that 
"technologies diffuse at different times, at different rates, and to
different extents in different places, and can be significantly
influenced by policies [@victor1993]". @meade2021modelling highlight that 
a variety of actors have a direct interest in gaining such knowledge 
including network equipment suppliers; network operators, regulatory 
and local authorities. 
These processes and their effects very a lot between scales: although
the diffusion of a new technology might not be optimal at a local level,
it might be beneficial from a global perspective as it could lead to 
faster diffusion to less advantaged places [@leibowicz2016representing]. 
Despite the spatial heterogeneity of such diffusion mechanisms and 
the policy relevance, there are very limited attempts in the literature 
to analyse the diffusion of new digital technologies at a detailed 
geographical level.

@@

Importantly, the data used here depicts the *active* engagement with the web as 
it contains all the commercial websites archived by the Internet Archive
^[[https://archive.org/](See https://archive.org/).]
that include a mention to at least one valid UK postcode in the web text. 
Websites are identified as commercial ones if they are part of the .co.uk second 
level domain (SLD). Also, the act of of creating a website is understood here as 
active engagement with the web vis-à-vis the more passive act of browsing the web 
or having an internet connection [@tranosuk]. Previous studies have focused mostly
on more passive notions of engaging with digital technologies such as internet 
adoption and internet speeds [e.g. @blank2018local; @riddlesden2014broadband].

@@

These ideas were firstly introduced by Torsten Hägerstrand and his thesis entitled 
'Innovation Diffusion as a Spatial Process' [@hagerstrand1968innovation]. 
Hägerstrand was the first one to identify diffusion as a geographical process. 
The starting point was the idea that diffusion is based on passing information 
through social networks, which themselves tend to be defined by geography. Hence, 
he identified the 'neighbourhood' effect of how information, and consequently, 
innovation diffuse. He used agricultural innovations to test and model his ideas 
using Monte Carlo simulations. Hägerstrand also incorporated the role of hierarchy 
and how some phenomena maybe firstly adopted in larger cities and then diffuse 
to second tier ones. This is a sequential instead of a simultaneous process, which 
resembles the 'lead-lag' spatial acceleration effect in market research 
[@bento2018time; @PERES201091]. Hägerstrand maybe more widely known for highlighting 
the role time plays in the diffusion of innovations: an early-pioneering 
period, a middle fast accelerating period and a final saturation one [@morrill2020spatial.

The temporal dimension was further explored by Everett Rogers and his seminal work 
on 'Diffusion of Innovations' [@rogers2010diffusion]. Rogers being a sociologist 
his work focused not on the diffusion of innovations over space and time, but 
instead on the adoption of new technologies and innovations by individuals and the 
individual mechanisms that drive the decisions behind adoption. He identified five 
groups of individuals regarding their adoption speed: innovators, early adopters, 
early majority, late majority and laggards. Key mechanism of diffusion and adoption 
is always communication and how knowledge is transferred within a social system. 
Therefore, all approaches agree on the S-shaped diffusion and cumulative adoption 
pattern [@grubler1990rise].

@@

<!-- previous studies -->
The remaining of this section reviews empirical studies which analysed the diffusion 
on new technologies over space and time. Although the spatial dimension is present
in most of the following studies, the level of spatial detail is almost always 
more coarse than the one adopted here. To begin with @beardsell1999spatial studied 
the evolution of the computer industry in $317$ US metro areas during the $1977-1992$ 
period using employment data. Their analysis indicated that the relative size distribution
holds for urban computer employment and also urban heterogeneity is essential in
explaining this distribution. @bednarz2020pulled focused on wind turbines and modelled
their spatial diffusion across $402$ German regions during $1970-2015$. Their key
finding is that local demand than local supply was the main driving factor.

At a global scale @perkins2005international analyzed if the diffusion rate of new
technologies is driven by a latecomer advantage and the engagement with the global
economy via foreign direct investments and trade. Their results illustrate that 
indeed latecomers and developing countries experience diffusion of new technologies 
more rapidly than early adopter and developed countries. At the same scale, 
@perkins2011internet explored whether the adoption of previous communication
technologies that is mail, telegrams and telephones was shaped by similar socioeconomic 
factors as the internet. Their results indicated common patterns regarding the 
drivers behind the adoption of different communication technologies.

@@

## Results {#sec3.1}

@@

-   Moran's I: for OA and LAD over time **TODO** add 0s

@@

-   S for LADs per firm and OA. **TODO**: OA per firm? fix firms over
    time.

@@

-   **TODo** check s_uk_firm_11

@@

## 3.2 Case study approach {#sec3.2}

# Digital technologies and spatial structure: a global view {#sec4}

# The impact of ICT on the US and the UK spatial structures {#sec5}

# Discussion and conclusions {#sec6}

>>>>

<!-- geographers lack of interest + motivation -->
Geographers were always interested in how new technologies and
innovations diffuse across space and time and, importantly, how such
spatio-temporal processes can be modelled. After all, diffusion together 
with invention and innovation are considered the pillars of 
technological change [@das2022diffusion]. The seminal contribution of
@hagerstrand1968innovation is illustrative of this early interest.
However, the torch of exploring and modelling such processes had been
passed to other disciplines such as economics, business studies and
sociology well before the 'cultural turn' of economic geography
[@perkins2005international]. A potential explanation of the lack of 
geographical studies exploring the diffusion of new and, more 
specifically for this paper, digital technologies across *both* space 
and time can be attributed to the scarcity of relevant and granular 
enough data. As @zook2022mapping highlight, digital activities are 
hardly ever captured in official data.

<!-- contribution -->
This paper offers such a contribution: a geographical study illustrating 
how a new technology that is the web diffused over space and time 
in the UK at a high level of spatial granularity during the $1996-2012$ 
period. It does so by employing a novel source of big data which 
captures the active engagement with web technologies during that period.
By addressing this empirical question this paper exemplifies how the 
combination of data sources which escape the traditional social science
domain and adequate research methods can offer new lenses to geographical
research regarding the understanding of technological diffusion.

<!-- motivation -->
The motivation for this paper lies in the fact that there are various
stakeholders who are interested in knowing how new digital technologies
diffused over space and time and use this knowledge to make predictions
regarding the diffusion of related *future* technologies.
As per @leibowicz2016representing, historical studies agree on the fact that 
"technologies diffuse at different times, at different rates, and to
different extents in different places, and can be significantly
influenced by policies [@victor1993]". @meade2021modelling highlight that 
a variety of actors have a direct interest in gaining such knowledge 
including network equipment suppliers; network operators, regulatory 
and local authorities. 
These processes and their effects very a lot across scales: although
the diffusion of a new technology might not be optimal at a local level,
it might be beneficial from a global perspective as it could lead to 
faster diffusion to less advantaged places [@leibowicz2016representing]. 
Despite the spatial heterogeneity of such diffusion mechanisms and 
the policy relevance, there are very limited attempts in the literature 
to analyse the diffusion of new digital technologies at a detailed 
geographical level.

@@

Importantly, the data used here depicts the *active* engagement with the web in 
the UK as it contains all the commercial websites archived by the Internet Archive
^[[https://archive.org/](See https://archive.org/).]
that include a mention to at least one valid UK postcode in the web text. 
Websites are identified as commercial ones if they are part of the .co.uk second 
level domain (SLD). Also, the act of of creating a website is understood here as 
active engagement with the web vis-à-vis the more passive act of browsing the web 
or having an internet connection [@tranosuk]. Previous studies have focused mostly
on more passive notions of engaging with digital technologies such as internet 
adoption and internet speeds [e.g. @blank2018local; @riddlesden2014broadband]. 
More details about the data and the data generation process can be found in Section
[XX](#sec3). **ADD DATA DESCRIPTION, WHY .CO.UK**

@@

These ideas were firstly introduced by Torsten Hägerstrand and his thesis entitled 
'Innovation Diffusion as a Spatial Process' [@hagerstrand1968innovation]. 
Hägerstrand was the first one to identify diffusion as a geographical process. 
The starting point was the idea that diffusion is based on passing information 
through social networks, which themselves tend to be defined by geography. Hence, 
he identified the 'neighbourhood' effect of how information, and consequently, 
innovation diffuse. He used agricultural innovations to test and model his ideas 
using Monte Carlo simulations. Hägerstrand also incorporated the role of hierarchy 
and how some phenomena maybe firstly adopted in larger cities and then diffuse 
to second tier ones. This is a sequential instead of a simultaneous process, which 
resembles the 'lead-lag' spatial acceleration effect in market research 
[@bento2018time; @PERES201091]. Hägerstrand is more widely known though for highlighting 
the role time plays in the diffusion of innovations: an early-pioneering period, 
a middle fast accelerating period and a final saturation one [@morrill2020spatial.

The temporal dimension was further explored by Everett Rogers and his seminal work 
on 'Diffusion of Innovations' [@rogers2010diffusion]. Rogers being a sociologist, 
his work focused not on the diffusion of innovations over space and time, but 
instead on the adoption of new technologies and innovations by individuals and the 
individual mechanisms that drive the decisions behind adoption. He identified five 
groups of individuals regarding their adoption speed: innovators, early adopters, 
early majority, late majority and laggards. Key mechanism of diffusion and adoption 
is always communication and how knowledge is transferred within a social system. 
Therefore, all approaches agree on the S-shaped diffusion and cumulative adoption 
pattern [@grubler1990rise].

@@

<!-- previous studies -->
The remaining of this section reviews empirical studies which analysed the diffusion 
on new technologies over space and time. Although the spatial dimension is present
in most of the following studies, the level of spatial detail is almost always 
more coarse than the one adopted here. To begin with @beardsell1999spatial studied 
the evolution of the computer industry in $317$ US metro areas during the $1977-1992$ 
period using employment data. Their analysis indicated that the relative size distribution
holds for urban computer employment and also urban heterogeneity is essential in
explaining this distribution. in a recent study, @bednarz2020pulled focused on 
wind turbines and modelled their spatial diffusion across $402$ German regions 
during $1970-2015$. Their key finding is that local demand than local supply was 
the main driving factor.

At a global scale @perkins2005international explored whther the diffusion rate of new
technologies is driven by a latecomer advantage and the engagement with the global
economy via foreign direct investments and trade. Their results illustrate that 
indeed latecomers and developing countries experience diffusion of new technologies 
more rapidly than early adopter and developed countries. At the same scale, 
@perkins2011internet explored whether the adoption of previous communication
technologies that is mail, telegrams and telephones was shaped by similar socioeconomic 
factors as the internet. Their results indicated common patterns regarding the 
drivers behind the adoption of different communication technologies.

@@

# Results {#sec3}

## Old ideas

@@

-   Moran's I: for OA and LAD over time

@@

-   S for LADs per firm and OA.

@@

-   s_uk_firm_11

@@

## New structure

1. S-shaped diffusion curves: S for LADs per firm. UK, fast/slow/examples

2. ranks: there is stability and movement

3. Neighbourhood effect: diffusion proceeds outwards from innovation
centers, first "hitting" nearby rather than far-away locations (Grubler
1990)

  -   Moran's I: for OA and LAD over time

  -   Website density regressions: for OA and LAD over time

4. Hierarchy effect: from main centers to secondary ones -- central places

5. RF

  - ideal: (i) train RF for all years and all (1) LADs and (2) OAs with CAST
  and report metrics. (ii) train for all years and per region for (1) LADs and (2)
  OAs, and predict to the other regions. Reports predictions as region similarities

# Discussion and conclusions {#sec4}

### 2023-11-24 paper/submission1/tranos2023.tex
CAT OA regions

\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\journal{Journal Name}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Digital economy in the UK: a multi-scalar story of the diffusion of web technologies},
  pdfauthor={Emmanouil Tranos},
  pdfkeywords={keyword1, keyword2},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{keyword}
    keyword1 \sep 
    keyword2
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, breakable, frame hidden, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, enhanced, interior hidden]}{\end{tcolorbox}}\fi

@@

Geographers were always interested in how new technologies and
innovations diffuse across space and, importantly, how such
spatio-temporal processes can be modelled. After all, diffusion together
with invention and innovation are considered the pillars of
technological change \citep{das2022diffusion}. The seminal contribution
of \citet{hagerstrand1968innovation} is illustrative of this early
interest. However, the torch of exploring and modelling such processes
had been passed to other disciplines such as economics, business studies
and sociology well before the `cultural turn' of economic geography
\citep{perkins2005international}. A potential explanation of the lack of
geographical studies exploring the diffusion of new and, more
specifically, digital technologies across \emph{both} space and time can
be attributed to the scarcity of relevant and granular enough data. As
\citet{zook2022mapping} highlight, digital activities are hardly ever
captured in official data.

This paper offers such a contribution: a geographical study illustrating
how a new technology that is the web diffused over space and time in the
UK at a high level of spatial granularity during the \(1996-2012\)
period. It does so by employing a novel source of big data which
captures the active engagement with web technologies during that period.
By addressing this empirical question this paper exemplifies how the
combination of data sources which escape the traditional social science
domain and adequate research methods can offer new contributions
regarding our understanding of how new technologies diffuse.

The motivation for this paper lies in the fact that there are various
stakeholders who are interested in knowing how new digital technologies
diffused over space and time and use this information to make
predictions regarding the diffusion of related \emph{future}
technologies. As per \citet{leibowicz2016representing}, historical
studies agree on the fact that ``technologies diffuse at different
times, at different rates, and to different extents in different places,
and can be significantly influenced by policies \citep{victor1993}''.
\citet{meade2021modelling} highlight that a variety of actors have a
direct interest in gaining such knowledge including network equipment
suppliers; network operators, regulatory and local authorities. These
processes and their effects very a lot between scales: although the
diffusion of a new technology might not be optimal at a local level, it
might be beneficial from a global perspective as it could lead to faster
diffusion to less advantaged places \citep{leibowicz2016representing}.
Despite the spatial heterogeneity of such diffusion mechanisms and the
policy relevance, there are very limited attempts in the literature to
analyse the diffusion of new digital technologies at a detailed
geographical level.

@@

This paper focuses on the diffusion of the web as new technology during
the \(1996-2012\) period. This was an exciting period for digital
technologies as it corresponds with the commercialisation of the
internet and, consequently, its almost universal adoption. The reader is
reminded that it was only in 1994 when Netscape Navigator was
introduced, a year before Microsoft's Internet Explorer \footnote{\url{https://www.theguardian.com/global/2015/mar/22/web-browser-came-back-haunt-microsoft}}.
Also, only \(9\) per cent of UK's households had access to the internet
in \(1998\) \citep{ons2018} and during that period that web included
mostly static webpages, there were no social media and web browsing was
happening exclusively from desktop PCs as there were no smartphones
\citep{tranosuk}. Hence, it is fair to say that the study period
captures the very early stages of the diffusion on a new technology that
is the web. This is key point in the lifecycle of a new technology.
Firstly, during this period new technologies are expensive, crude and
imperfect \citep{rosenberg1994exploring, wilson201281}. A simple
comparison between Web 1.0 and Web 3.0 applications clearly illustrates
this \citep{tranos2020social}. During this period the performance of a
new technology is the main attraction and not the cost to access and use
it \citep{wilson2011lessons} There is a broader theoretical discussion
in the literature regarding the motivation behind early adoption. As
summarised by \citet{perkins2005international}, on the one hand,
epidemic models highlight the role of interpersonal contacts as a way
for new technologies to diffuse. On the other hand, economic models
underline the importance of heterogeneity. Different firms have
different structures and business plans, which define the potential
economic returns of the adoption of a new technology and, therefore, the
choice to adopt a new technology becomes an individual option. From a
broader and evolutionary perspective, initial conditions are essential
for the creation and evolution of path-dependent technological
development trajectories \citep{neffke2011regions, simmie2014new}. This
argument is even more relevant when the focus is on Information and
Communication Technologies (ICT) because of the commonly found lag
between investment and economic returns as reflected in the Solow
paradox \citep{acemoglu2014return, brynjolfsson2018artificial}.

Importantly, the data used here depicts the \emph{active} engagement
with the web as it contains all the commercial websites archived by the
Internet Archive \footnote{\href{See\%20https://archive.org/}{https://archive.org/}.}
that include a mention to at least one valid UK postcode in the web
text. Websites are identified as commercial ones if they are part of the
.co.uk second level domain (SLD). Also, the act of of creating a website
is understood here as active engagement with the web vis-à-vis the more
passive act of browsing the web or having an internet connection
\citep{tranosuk}. Previous studies have focused mostly on more passive
notions of engaging with digital technologies such as internet adoption
and internet speeds
\citep[e.g.][]{blank2018local, riddlesden2014broadband}.

>>>>

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

\makeatletter
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\makeatletter
\@ifpackageloaded{tcolorbox}{}{\usepackage[skins,breakable]{tcolorbox}}
\makeatother
\makeatletter
\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\makeatother
\journal{Journal Name}
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage[]{natbib}
\bibliographystyle{elsarticle-harv}
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Digital economy in the UK: a multi-scalar story of the diffusion of web technologies},
  pdfauthor={Emmanouil Tranos},
  pdfkeywords={keyword1, keyword2},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

@@

\begin{keyword}
    keyword1 \sep 
    keyword2
\end{keyword}
\end{frontmatter}
    \ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, breakable, frame hidden, interior hidden, sharp corners, enhanced, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

Geographers were always interested in how new technologies and
innovations diffuse across space and time and, importantly, how such
spatio-temporal processes can be modelled. After all, diffusion together
with invention and innovation are considered the pillars of
technological change \citep{das2022diffusion}. The seminal contribution
of \citet{hagerstrand1968innovation} is illustrative of this early
interest. However, the torch of exploring and modelling such processes
had been passed to other disciplines such as economics, business studies
and sociology well before the `cultural turn' of economic geography
\citep{perkins2005international}. A potential explanation of the lack of
geographical studies exploring the diffusion of new and, more
specifically for this paper, digital technologies across \emph{both}
space and time can be attributed to the scarcity of relevant and
granular enough data. As \citet{zook2022mapping} highlight, digital
activities are hardly ever captured in official data.

This paper offers such a contribution: a geographical study illustrating
how a new technology that is the web diffused over space and time in the
UK at a high level of spatial granularity during the \(1996-2012\)
period. It does so by employing a novel source of big data which
captures the active engagement with web technologies during that period.
By addressing this empirical question this paper exemplifies how the
combination of data sources which escape the traditional social science
domain and adequate research methods can offer new lenses to
geographical research regarding the understanding of technological
diffusion.

The motivation for this paper lies in the fact that there are various
stakeholders who are interested in knowing how new digital technologies
diffused over space and time and use this knowledge to make predictions
regarding the diffusion of related \emph{future} technologies. As per
\citet{leibowicz2016representing}, historical studies agree on the fact
that ``technologies diffuse at different times, at different rates, and
to different extents in different places, and can be significantly
influenced by policies \citep{victor1993}''. \citet{meade2021modelling}
highlight that a variety of actors have a direct interest in gaining
such knowledge including network equipment suppliers; network operators,
regulatory and local authorities. These processes and their effects very
a lot across scales: although the diffusion of a new technology might
not be optimal at a local level, it might be beneficial from a global
perspective as it could lead to faster diffusion to less advantaged
places \citep{leibowicz2016representing}. Despite the spatial
heterogeneity of such diffusion mechanisms and the policy relevance,
there are very limited attempts in the literature to analyse the
diffusion of new digital technologies at a detailed geographical level.

@@

This paper focuses on the diffusion of the web as new technology during
the \(1996-2012\) period. This was an exciting period for digital
technologies as it corresponds with the commercialisation of the
internet and, consequently, its almost universal adoption. The reader is
reminded that it was only in 1994 when Netscape Navigator was
introduced, a year before Microsoft's Internet Explorer \footnote{\url{https://www.theguardian.com/global/2015/mar/22/web-browser-came-back-haunt-microsoft}}.
Also, only \(9\) per cent of UK's households had access to the internet
in \(1998\) \citep{ons2018}, the web included mostly static webpages,
there were no social media and web browsing was happening exclusively
from desktop PCs as there were no smartphones \citep{tranosuk}. Hence,
it is fair to say that the study period captures the very early stages
of the diffusion on a new technology that is the web. This is key point
in the lifecycle of a new technology. Firstly, during this period new
technologies are expensive, crude and imperfect
\citep{rosenberg1994exploring, wilson201281}. A simple comparison
between Web 1.0 and Web 3.0 applications clearly illustrates this
\citep{tranos2020social}. During this period the performance of a new
technology is the main attraction and not the cost to access and use it
\citep{wilson2011lessons} There is a broader theoretical discussion in
the literature regarding the motivation behind early adoption. As
summarised by \citet{perkins2005international}, on the one hand,
epidemic models highlight the role of interpersonal contacts as a way
for new technologies to diffuse. On the other hand, economic models
underline the importance of heterogeneity. Different firms have
different structures and business plans, which define the potential
economic returns of the adoption of a new technology and, therefore, the
choice to adopt a new technology becomes an individual option. From a
broader and evolutionary perspective, initial conditions are essential
for the creation and evolution of path-dependent technological
development trajectories \citep{neffke2011regions, simmie2014new}. This
argument is even more relevant when the focus is on Information and
Communication Technologies (ICT) because of the commonly found lag
between investment and economic returns as reflected in the Solow
paradox \citep{acemoglu2014return, brynjolfsson2018artificial}.

Importantly, the data used here depicts the \emph{active} engagement
with the web in the UK as it contains all the commercial websites
archived by the Internet Archive \footnote{\href{See\%20https://archive.org/}{https://archive.org/}.}
that include a mention to at least one valid UK postcode in the web
text. Websites are identified as commercial ones if they are part of the
.co.uk second level domain (SLD). Also, the act of of creating a website
is understood here as active engagement with the web vis-à-vis the more
passive act of browsing the web or having an internet connection
\citep{tranosuk}. Previous studies have focused mostly on more passive
notions of engaging with digital technologies such as internet adoption
and internet speeds
\citep[e.g.][]{blank2018local, riddlesden2014broadband}. More details
about the data and the data generation process can be found in Section
\protect\hyperlink{sec3}{XX}. \textbf{ADD DATA DESCRIPTION, WHY .CO.UK}
